---
title: "Example used in manuscript"
author: "Xavier Didelot"
date: '`r Sys.Date()`'
output:
  pdf_document: default
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Example used in manuscript}
  %\usepackage[utf8]{inputenc}
---

This is the code to reproduce an example shown in the manuscript.
The main point is to show that ignoring changes in the local population size
leads to false positives.

# Initialisation

```{r}
library(DetectImports)
library(ape)
set.seed(2)
```

# Simulation with no import but varying population size

```{r}
set.seed(0)
Ne=function(t){if (t>2020.5) return(1) else return(0.2)}
tree=simCoal(seq(2020,2021,length.out = 100),Ne)
plotBoth(tree,Ne)
```

# Inference assuming constant population size

Let's detect the imports under the constant population size model:

```{r}
res=detectImports(tree,constant=T,verbose=F,seed=0)
plot(res)
```

# Inference under full model

Let's detect the imports under the full model:

```{r}
res2=detectImports(tree,verbose=F,seed=0)
plot(res2)
```

# Combine into a figure to include in paper

```{r}
pdf('/tmp/figFalsePos.pdf',5,7)
par(mfrow=c(3,1),mar = c(3,4,3,4))#,xpd=NA)
plotImports(tree,tree$imports,xaxs='i')
plot(res,bty='l',xlim=c(2021-max(dist.nodes(tree)[Ntip(tree)+1,]),2021),xaxs='i',xlab='')
plot(res2,bty='l',xlim=c(2021-max(dist.nodes(tree)[Ntip(tree)+1,]),2021),xaxs='i',xlab='')
par(xpd=NA)
x=2021-max(dist.nodes(tree)[Ntip(tree)+1,])*1.1
y=max(tree$stats[,'coalint'],na.rm = T)
text(x,y*4.3,'A',cex=2)
text(x,y*2.75,'B',cex=2)
text(x,y*1.1,'C',cex=2)
dev.off()
system('open /tmp/figFalsePos.pdf')
```


# Get some pvals for the manuscript text
```{r}
min(res$pvals,na.rm = T)#lowest pval according to res
min(res2$pvals,na.rm = T)#lowest pval according to res2
```
