---
title: "Demonstration of DetectImports"
author: "Xavier Didelot"
date: '`r Sys.Date()`'
output:
  pdf_document: default
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Demosntration of DetectImports}
  %\usepackage[utf8]{inputenc}
---

# Initialisation

```{r}
library(DetectImports)
library(ape)
set.seed(0)
```

# Simulation with constant population size 

For initial simplicity, consider that the local population has constant effective population size $N_e$.

```{r}
Ne=function(t){return(5)}
tree=simCoal(seq(2000,2020,0.05),Ne)
plotBoth(tree,Ne)
```

# Coalescent intervals

We consider the leaves $i=1..n$ in increasing order of date. Let $t_i$ denote the date of leaf $i$, we have $i<j \implies t_i<t_j$. The $j$-th leaf has to coalesce with the genealogy formed by the $(j-1)$ first leaves. Let $A_i$ denote the sum of branch lengths in the genealogy formed so far between the time of leaf $i$ and the time where it coalesces with an ancestor of a previously considered leaf. We call $A_i$ the ``coalescent interval''. Let's plot $A_i$ on the y-axis against $t_i$ on the x-axis:

```{r}
m=keyStats(tree)
dates=m[2:Ntip(tree),'dates']
coalints=m[2:Ntip(tree),'coalint']
plot(dates,coalints,
     xlab='Sampling dates',ylab='Coalescent intervals')
```

# Detection of imports

Imports are likely to have $A_i$ greater than expected if transmission happened only locally.
In the case where the location population size is constant equal to $N_e$, 
$A_i$ is exponentially distributed with mean $N_e$. 
So we can detect imports by looking for coalescent intervals that are greater than this. 
Of course here there are no imports:

```{r}
pvals=test0(dates,coalints,Ne(0))
summary(pvals)
```

# Simulation with varying population size

Now relax so that the effective population size is not constant but is a continuous function $Ne(t)$ of time.  

```{r}
Ne=function(t){return(sin(t)+1.1)}
tree=simCoal(seq(2000,2020,0.05),Ne)
plotBoth(tree,Ne)
```

# Analysis of tree with varying population size

The coalescent intervals $A_i$ are no longer exponential and no longer iid. But we can still do the same plot as before:

```{r}
m=keyStats(tree)
dates=m[2:Ntip(tree),'dates']
coalint=m[2:Ntip(tree),'coalint']
plot(dates,coalint,
     xlab='Sampling dates',ylab='Coalescent interval')
```

# Detection of imports

Outliers in the distribution above are likely imports. 

We would need a non-parametric test to detect outliers (has to be non-parametric because the distribution $p(A_i|t_i)$ is unknown and complex in the general case). Or perhaps can we use semi-parametric test in which the distribution is assumed exponential with time-changing mean $\hat N_e(t)$ estimated from non-outlier points. This is not quite correct since the distribution is not exponential, but matching the mean might be enough.

# Simulation with imports

Since only the local population was simulated without any imports, in the cases above there were no imports. But we could do similar simulations with imports using the structured coalescent, with two demes (local and global), only migration from global to local, bigger population in global, and a varying rate of migration from global to local over time. 
One option to do these simulations would be to use the`scrm` and/or the `coala` R packages.


