---
title: "Demonstration of DetectImports"
author: "Xavier Didelot"
date: '`r Sys.Date()`'
output:
  pdf_document: default
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Demosntration of DetectImports}
  %\usepackage[utf8]{inputenc}
---

# Initialisation

```{r}
library(DetectImports)
library(ape)
set.seed(0)
```

# Simulation with constant population size 

For initial simplicity, consider that the local population has constant effective population size $N_e$.

```{r}
Ne=function(t){return(5)}
tree=simCoal(seq(2000,2020,0.05),Ne)
plotBoth(tree,Ne)
```

# Coalescent intervals

We consider the leaves $i=1..n$ in increasing order of date. Let $t_i$ denote the date of leaf $i$, we have $i<j \implies t_i<t_j$. The $j$-th leaf has to coalesce with the genealogy formed by the $(j-1)$ first leaves. Let $A_i$ denote the sum of branch lengths in the genealogy formed so far between the time of leaf $i$ and the time where it coalesces with an ancestor of a previously considered leaf. We call $A_i$ the ``coalescent interval''. Let's plot $A_i$ on the y-axis against $t_i$ on the x-axis:

```{r}
plotCoalInt(tree)
```

# Detection of imports

Imports are likely to have $A_i$ greater than expected if transmission happened only locally.
In the case where the location population size is constant equal to $N_e$, 
$A_i$ is exponentially distributed with mean $N_e$. 
So we can detect imports by looking for coalescent intervals that are greater than this. 
Since $N_e$ is unknown, we need to estimate it from the data.
A simple estimator would be to take the mean of the $A_i$, but since we suspect that some $A_i$
are affected by imports, it is preferable to use an estimator based on the median:
$$\hat N_e=\frac{\mathrm{median}(A_{1..n})}{\mathrm{ln}(2)}$$
Of course here there are no imports:

```{r}
pvals=test0(tree)
```

# Simulation with varying population size

Now we consider that the effective population size is not constant but is a continuous function $N_e(t)$ of time.

```{r}
set.seed(0)
Ne=function(t){if (t>2010|t<2005) return(2) else return(10)}
tree=simCoal(seq(2000,2020,0.05),Ne)
plotBoth(tree,Ne)
```

# Coalescent intervals

The coalescent intervals $A_i$ are no longer exponential and no longer iid as in the case of a constant population size. But we can still compute them and plot them as before:

```{r}
plotCoalInt(tree)
```

# Bad test

If we apply the same test as before, we will get some false positives in the detection of imports:

```{r}
pvals=test0(tree)
```

# Detection of imports

Outliers in the distribution of $A_i$ against $t_i$ are likely imports. 
We could use a non-parametric test to detect outliers, this would be good since 
the distribution $p(A_i|t_i)$ is unknown and complex in the general case. 

Alternatively we could use semi-parametric test in which the distribution is assumed exponential with time-changing mean $\hat N_e(t)$ estimated from non-outlier points. This is not quite correct since the distribution is not exponential, but matching the mean might be enough to give a robust test.

As a first attempt, let's consider the estimator:
$$\hat N_e(t_i)=\frac{\mathrm{median}(A_{j\neq i\mathrm{~with~}|t_i-t_j|<\epsilon})}{\mathrm{ln}(2)}$$

Let's try this:
```{r}
pvals=test1(tree,epsilon=1)
```

# Simulation with imports

Since only the local population was simulated in the examples above there were no imports. But we could do similar simulations with imports using the structured coalescent, with two demes (local and global), only migration from global to local, bigger population in global, and a varying rate of migration from global to local over time. 
This would allow us to show that the tests are not only specific but also sensitive.
One option to do these simulations would be to use the `scrm` and/or the `coala` R packages.


