---
title: "Example used in manuscript"
author: "Xavier Didelot"
date: '`r Sys.Date()`'
output:
  pdf_document: default
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Example used in manuscript}
  %\usepackage[utf8]{inputenc}
---

This is the code to reproduce an example shown in the manuscript.
The main point is to show that ignoring changes in the local population size
leads to false negatives.

# Initialisation

```{r}
library(DetectImports)
library(ape)
set.seed(1)
```

# Simulation with a single import

Let's consider that the local population started in 2020 and a single import occurred in 2020.5.
500 genomes were sampled between 2020 and 2021.

```{r}
tree=simImports(localPopStart=2020,importDates=c(2020.25),
                samplingStartDate=2020,samplingEndDate=2021,samplingNumber=500,globalNeg = 0.01)
plot(tree,show.tip.label = F)
axisPhylo(backward = F)
```

# Inference assuming constant population size

Let's detect the imports under the constant population size model:

```{r}
res=detectImports(tree,constant=T,verbose=F,seed=0)
plot(res)
```

# Inference under full model

Let's detect the imports under the full model:

```{r}
res2=detectImports(tree,verbose=F,seed=0)
plot(res2,xlab='')
```

# Combine into a figure to include in paper

```{r}
pdf('/tmp/figFalseNeg.pdf',5,7)
par(mfrow=c(3,1),mar = c(3,4,3,4),xpd=NA)
plotImports(tree,tree$imports,xaxs='i',show.axis = F,colorDescendants = 'black')
labs=c('Jan','Mar','May','Jul','Sep','Nov','Jan')
axis(1,at = c(-1,2020+seq(0,1,1/6)-tree$root.time),labels=c('',labs),xaxs='i')
plot(res,bty='l',xlim=c(2021-max(dist.nodes(tree)[Ntip(tree)+1,]),2021),xaxs='i',xlab='',xaxt='n')
axis(1,at = 2020+seq(0,1,1/6),labels=labs,xaxs='i')
plot(res2,bty='l',xlim=c(2021-max(dist.nodes(tree)[Ntip(tree)+1,]),2021),xaxs='i',xlab='',xaxt='n')
axis(1,at = 2020+seq(0,1,1/6),labels=labs,xaxs='i')
par(xpd=NA)
x=2021-max(dist.nodes(tree)[Ntip(tree)+1,])*1.1
y=max(tree$stats[,'coalint'],na.rm = T)
text(x,y*4.3,'A',cex=2)
text(x,y*2.75,'B',cex=2)
text(x,y*1.1,'C',cex=2)
dev.off()
system('open /tmp/figFalseNeg.pdf')
```

# Get some pvals for the manuscript text
```{r}
res$pvals[tree$imports]#pval of correct import according to res
res2$pvals[tree$imports]#pval of correct import according to res2
min(res$pvals[-tree$imports],na.rm = T)#lowest pval for incorrect import according to res
min(res2$pvals[-tree$imports],na.rm = T)#lowest pval for incorrect import according to res2
```
