---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
set.seed(0)
```

# DetectImports

The goal of DetectImports is to detect importations in a dated phylogeny of local samples only, ie no samples from external locations (otherwise it becomes more like a phylogeography type of inference). I would like to make as few assumptions as possible about what is going on outside of the local population. If there were only local transmission, we can describe what phylogenies are expected to look like under a coalescent model with varying population size. We can think about this as the null distribution, and calculate for each sample the probability of importation based on how unlikely the observation would be under a model of local transmission only.

## Installation

You can install DetectImports from github with:

```{r gh-installation, eval = FALSE}
devtools::install_github("xavierdidelot/DetectImports")
```

The package can then be loaded using:
```{r}
library(DetectImports)
```

## Example

Cf vignette.
